Galactic Chemical Evolution Models Favor an Extended Type Ia Supernova Delay-Time Distribution

### 2024-03-21 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of the Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and \aFe in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in %one-zone and 
    multi-zone chemical evolution models including a prescription for radial stellar migration.
    % using the Versatile Integrator for Chemical Evolution (\vice). \vice includes a prescription for radial stellar migration in multi-zone models. 
    We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the final data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the SN Ia DTD has the largest effect on the \aFe distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower \aFe ratio, whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher \aFe. While the DTD alone cannot explain the observed bimodality in the \aFe distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence. Our model results favor a DTD with longer delay times and fewer prompt SNe Ia than the fiducial $t^{-1}$ power law.
\end{abstract}

@@

% [What is the delay time distribution]
The {\it delay time distribution} (DTD) refers to the rate of SN Ia events per unit mass of star formation as a function of stellar population age
% The rate of SNe Ia which explode as a function of time after the formation of a hypothetical unit-mass single stellar population is known as the delay-time distribution 
\citep[for a review, see Section 3.5 of][]{Maoz2014-Review}.
When the DTD is convolved with the Galactic star formation rate (SFR), it yields the overall SN Ia rate. 
The quantitative details of the relationship between \aFe and [Fe/H] are set by the DTD, and as such it is a key parameter in GCE models.
However, the DTD remains poorly constrained because it reflects the detailed evolution of the SN Ia progenitor systems, so different models for the progenitors of SNe Ia will naturally predict different forms for the DTD.

@@

% [Observational DTDs]
Surveys of SNe Ia 
% have provided the primary observational constraints on the DTD. This can be done 
can constrain the DTD
by comparing the observed rate of SNe Ia to their host galaxy parameters \citep[e.g.,][]{Mannucci2005-SNRate,Heringer2019-FieldGalaxyDTD} or inferred star formation histories \citep[SFHs; e.g.,][]{Maoz2012-SloanIIDTD}, measuring SN Ia rates in galaxy clusters \citep[e.g.,][]{Maoz2010-ClusterDTD}, or comparing the volumetric SN Ia rate to the cosmic SFH as a whole \citep[e.g.,][]{Graur2014-VolumetricSNIaRates,Strolger2020-ExponentialDTD}. 
Early studies, which had limited sample sizes, produced unimodal \citep{Strolger2004-SNIaProgenitors} or bimodal \citep{Mannucci2006-TwoPopulations} DTDs where the majority of SNe Ia explode within a relatively narrow range of delay times.
More recent studies have recovered broader DTD functions, with many converging on a declining power-law of $\sim t^{-1}$ \citep[e.g.,][]{Maoz2017-CosmicDTD,Castrillo2021-DTD,Wiseman2021-DESRates}, though there is some evidence for a steeper slope in galaxy clusters \citep{Maoz2017-CosmicDTD,FriedmannMaoz2018-ClusterDTD}.
It is especially difficult to constrain the DTD for short delay times \citep{MaozMannucci2012-SNeIaReview} because of the need for SN Ia rates at long lookback times and uncertainties in the age estimates of stellar populations.

@@

The uncertainties in the SN Ia DTD propagate into GCE models. In principle, the observed chemical abundance patterns should therefore contain information about the SN Ia DTD, and by extension their progenitors.
A few studies have compared different DTDs in one-zone chemical evolution models, but their comparisons to abundance data are limited to the local solar neighborhood \citep[e.g.,][]{Andrews2017-ChemicalEvolution,Palicio2023-AnalyticDTD}. 
[Talk about \citep{Matteucci2009-DTDModels} multi-zone models].
\citet{Poulhazan2018-PrecisionPollution} compared six DTDs in a cosmological smoothed-particle hydrodynamics simulation. They found that models with a prompt SN Ia component produced the highest abundance ratios and the narrowest \aFe distributions at late times, although their simulation was not tuned to reproduce the parameters of the Milky Way, and they did not compare to observed chemical abundances. 
The current era of large spectroscopic surveys such as the Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE;][]{Majewski2017-APOGEE} and the forthcoming Milky Way Mapper \citep{Kollmeier2017-SDSS-V} has made abundances across the Milky Way disk available for comparison to more sophisticated GCE models.

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit some parameters that are unchanged from \citetalias{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone model parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:multizone-results} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:multizone-results} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\uptau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration distance \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\uptau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:multizone-results} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:multizone-results} & Maximum simulation time \\
        n               & 8         & \ref{sec:multizone-results} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:multizone-results} & Maximum radius of star formation \\
        M_{g,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        \dot M_r    & continuous    & \ref{sec:multizone-results} & Recycling rate \citep[][Equation 2]{JohnsonWeinberg2020-Starbursts} \\
        \hline
        % DTD
        R_{\rm Ia}(t)   & Equation \ref{eq:dtd-function}    & \ref{sec:dtd-models}  & Delay time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        N_{\rm Ia}/M_\star  & $2.2\times10^{-3}$ M$_\odot^{-1}$ & \ref{sec:yields}  & SNe Ia per unit mass of stars formed \citep{MaozMannucci2012-SNeIaReview} \\
        \hline
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & \ref{sec:yields}  & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & \ref{sec:yields}  & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & \ref{sec:yields}  & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & \ref{sec:yields}  & SN Ia yield of Fe \\
        \hline
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the inside-out SFR \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the late-burst SFR \\
        \uptau_{\rm rise}             & 2 Gyr     & \ref{sec:sfh} & SFR rise timescale for inside-out and early-burst models \\
        \uptau_{\rm EB}(t)          & Equation \ref{eq:earlyburst-taustar}  & \ref{sec:sfh}   & Time-dependence of the early-burst SFE timescale \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the early-burst IFR \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the two-infall IFR \\
        \hline
        % One-zone parameters
        \uptau_\star                    & 2 Gyr & \ref{sec:onezone-results} & SFE timescale in one-zone models \\
        \eta(R_{\rm gal}=8\,{\rm kpc})  & 2.15  & \ref{sec:onezone-results} & Outflow mass-loading factor at the solar annulus \\
        \uptau_{\rm sfh}(R_{\rm gal}=8\,{\rm kpc})    & 15.1 Gyr  & \ref{sec:sfh} & SFH timescale at the solar annulus \\
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

\subsection{Delay Time Distributions}
\label{sec:dtd-models}

@@

\begin{figure}
    \centering
    \includegraphics{figures/delay_time_distributions.pdf}
    \caption{Selection of models for the SN Ia delay time distribution (DTD) used in this paper. All functions are normalized such that $f_{\rm Ia}(t=1\,{\rm Gyr})=1$. The black squares represent the DTD recovered for the SDSS-II sample of SNe Ia by \citet{Maoz2012-SloanIIDTD} at the same scale as the DTDs. The horizontal and vertical error bars indicate the time range and 1$\sigma$ uncertainties, respectively, of each DTD measurement. The colored circles along the horizontal axis indicate the median delay time for each model.}
    \label{fig:dtds}
    \script{delay_time_distributions.py}
\end{figure}

% \begin{deluxetable*}{lcll}
% \tablecaption{Summary of SN Ia DTDs explored in this paper.\label{tab:dtds}}
% \tablehead{
% \colhead{Model} & \colhead{Eq.} & \colhead{Parameters} & \colhead{Similar to}
% }
% \startdata
\begin{table*}
    \centering
    \caption{Summary of SN Ia delay time distributions (DTDs) explored in this paper.}
    \label{tab:dtds}
    \begin{tabular}{lcll}
        \hline\hline
        \multicolumn{1}{c}{Model} & Eq. & \multicolumn{1}{c}{Parameters} & \multicolumn{1}{c}{Similar to} \\
        \hline
        Two-population  & \ref{eq:prompt-dtd}   & $t_{\rm max}=0.05$ Gyr, $\sigma=0.015$ Gyr, & \citet{Mannucci2006-TwoPopulations} \\
                        &                       & $\uptau_{\rm Ia}=3$ Gyr & \\
        Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.4$                 & \citet[][cluster]{Maoz2017-CosmicDTD}; 
                                                      \citet{Heringer2019-FieldGalaxyDTD}       \\
        Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.1$                 & \citet[][field]{Maoz2017-CosmicDTD}; 
                                                      \citet{Wiseman2021-DESRates}              \\
        Exponential & \ref{eq:exponential-dtd}  & $\uptau_{\rm Ia}=1.5$ Gyr    & \citet[][SD]{Greggio2005-AnalyticalRates};
                                                                      \citet{Schonrich2009-RadialMixing};       \\
                    &                           &                   & \citet{Weinberg2017-ChemicalEquilibrium}  \\
        Exponential & \ref{eq:exponential-dtd}  & $\uptau_{\rm Ia}=3$ Gyr   & --- \\
        Plateau     & \ref{eq:plateau-dtd}  & $W=0.3$ Gyr, $\alpha=-1.1$    & \citet[][CLOSE DD]{Greggio2005-AnalyticalRates} \\
        Plateau     & \ref{eq:plateau-dtd}  & $W=1$ Gyr, $\alpha=-1.1$      & \citet[][WIDE DD]{Greggio2005-AnalyticalRates} \\
        Triple-system   & \ref{eq:triple-dtd}   & $f_{\rm init}=0.05f_{\rm peak}$, $t_{\rm rise}=0.5$ Gyr, & \citet{Rajamuthukumar2023-TripleEvolution} \\
                        &                       & $W=0.5$ Gyr, $\alpha=-1.1$ & \\
        \hline
    \end{tabular}
\end{table*}
% \enddata
% \end{deluxetable*}

@@

% [Intro paragraph, motivation for running one-zone models]
Before running the full multi-zone models, it is useful to understand the effects of the DTD in more idealized conditions. A one-zone model assumes the entire gas reservoir is instantaneously mixed, removing all spatial dependence. This limits the ability to compare to observations across the disk, but it eliminates the complicating factor of stellar migration and better isolates the effects of the nucleosynthesis prescription. In this section, we compare the results from one-zone models which examine various parameters of the DTD while keeping other parameters fixed.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/feh_df_comparison.pdf}
    \caption{Distributions of [Fe/H] from multi-zone models with various models for the SFH and DTD. Each row presents distributions of stars within a range of absolute midplane distance: $1\leq|z|<2$ kpc (\textit{top}), $0.5\leq|z|<1$ kpc (\textit{middle}), and $0\leq|z|<0.5$ kpc (\textit{bottom}). Within each panel, curves of different color represent the distributions of stars binned by Galactocentric radius $R_{\rm gal}$, from $3\leq R_{\rm gal}<5$ kpc (yellow) to $13\leq R_{\rm gal}<15$ kpc (blue). Each distribution is normalized so the area under the curve is 1, and the vertical scale is consistent across each row. All distributions are convolved with observational uncertainties in APOGEE DR17 (see Table \ref{tab:sample}) and smoothed with a box-car width of 0.2 dex. 
    \textit{Left columns:} comparison between the inside-out and two-infall SFHs; both assume the exponential ($\uptau_{\rm Ia}=1.5$ Gyr) DTD. 
    \textit{Center column:} the distributions from APOGEE DR17 for reference, binned and smoothed similarly.
    \textit{Right columns:} comparison between the power-law ($\alpha=-1.4$) and exponential ($\uptau_{\rm Ia}=3$ Gyr) DTDs with the inside-out SFH. [Add: noticeable but subtle change in inner-galaxy peak and tail between DTDs]}
    \label{fig:feh-df-comparison}
    \script{feh_df_comparison.py}
\end{figure*}

@@

% The choice of DTD and SFH has a larger effect on the inner regions of the Galaxy than the outer. Inner regions have a higher SFR and more stars are formed at early times than late, so differences in the SN Ia rate have a more pronounced effect on the MDF. For all SFHs except two-infall, the outer Galaxy has a longer SFR timescale than the inner Galaxy, so the MDF in the outer Galaxy is relatively unaffected by the fraction of prompt SNe Ia.
The inner Galaxy MDF is more sensitive to the choice of DTD than the outer Galaxy. The SFH peaks earlier and declines more sharply due to the inside-out formation of the disk. Consequently, SNe Ia often explode when the gas supply is significantly lower than when the progenitors formed. This so-called ``gas-starved ISM'' effect drives a faster increase in metallicity \citep[see analytic demonstration in][]{Weinberg2017-ChemicalEquilibrium}, which ultimately lowers the number of low-metallicity stars. The more extended the DTD< the stronger the effect. The outer disk is less affected by the choice of DTD, though, due to the more extended SFH.

While there are some {\it quantitative} differences in how the shape of the MDF varies with Galactic region, the {\it qualitative} trends are unaffected by the choice of model SFH or DTD. These trends are primarily driven by chemical equilibrium, abundance gradients, and radial migration \citepalias[][see their Section 3.2 for further discussion]{Johnson2021-Migration}.

@@

The model combination with the best (lowest) score for the MDF is the two-infall SFH with the exponential ($\uptau_{\rm Ia}=3$ Gyr) DTD. The choice of SFH has a larger effect on the overall score than the DTD, and the best-performing SFH is the two-infall model, followed closely by the early-burst. However, the difference between the best-scoring model and the worst (late-burst SFH with triple-evolution DTD) is fairly small. The quantitative agreement between the predicted and observed MDFs is primarily affected by the assumption of chemical equilibrium, the abundance gradient, and radial migration (see discussion in section 3.2 of \citetalias{Johnson2021-Migration}).

@@

A striking feature of the [O/Fe] distribution in the Milky Way is the distinct separation into two components, the high- and low-$\alpha$ sequences [references]. The relative prominence of the two peaks varies with $R_{\rm gal}$ and $|z|$, as shown by the right-hand panels of Figures \ref{fig:ofe-df-sfh} and \ref{fig:ofe-df-dtd}. A crucial feature of this bimodality, which is not apparent in the previous figures, is the presence of both sequences at the same [Fe/H]. Figure \ref{fig:ofe-bimodality} compares the [O/Fe] distributions in the solar annulus ($7\leq R_{\rm gal}<9$ kpc and $0\leq|z|<2$ kpc) in two bins of [Fe/H] for select model outputs and APOGEE data. The APOGEE distributions in the bottom-right panel (j) show that the high-$\alpha$ mode is more prominent at lower [Fe/H] and vice versa, but the distributions in both bins are clearly bimodal. Moreover, the ``trough'' at ${\rm [O/Fe]}\approx0.2$ is consistent in each bin.

@@

\begin{itemize}

>>>>

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of the Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation, and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and \aFe in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in %one-zone and 
    multi-zone chemical evolution models which include radial stellar migration.
    % using the Versatile Integrator for Chemical Evolution (\vice). \vice includes a prescription for radial stellar migration in multi-zone models. 
    We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the final data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the DTD has the largest effect on the \aFe distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower \aFe ratio.
    % whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher \aFe.
    While the DTD alone cannot explain the observed bimodality in the \aFe distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence. Our model results favor an extended DTD with fewer prompt SNe Ia than the fiducial $t^{-1}$ power law.
\end{abstract}

@@

% [What is the delay-time distribution]
The {\it delay-time distribution} (DTD) refers to the rate of SN Ia events per unit mass of star formation as a function of stellar population age
% The rate of SNe Ia which explode as a function of time after the formation of a hypothetical unit-mass single stellar population is known as the delay-time distribution 
\citep[for a review, see Section 3.5 of][]{Maoz2014-Review}.
When the DTD is convolved with the Galactic star formation rate (SFR), it yields the overall SN Ia rate. 
The quantitative details of the relationship between \aFe and [Fe/H] are set by the DTD, and as such it is a key parameter in GCE models.
However, the DTD remains poorly constrained because it reflects the detailed evolution of the SN Ia progenitor systems, so different models for the progenitors of SNe Ia will naturally predict different forms for the DTD.

@@

% [Observational DTDs]
Surveys of SNe Ia can constrain the DTD by comparing the observed rate of SNe Ia to their host galaxy parameters \citep[e.g.,][]{Mannucci2005-SNRate,Heringer2019-FieldGalaxyDTD} or inferred star formation histories \citep[SFHs; e.g.,][]{Maoz2012-SloanIIDTD}, measuring SN Ia rates in galaxy clusters \citep[e.g.,][]{Maoz2010-ClusterDTD}, or comparing the volumetric SN Ia rate to the cosmic SFH as a whole \citep[e.g.,][]{Graur2014-VolumetricSNIaRates,Strolger2020-ExponentialDTD}. 
Early studies, which had limited sample sizes, produced unimodal \citep{Strolger2004-SNIaProgenitors} or bimodal \citep{Mannucci2006-TwoPopulations} DTDs where the majority of SNe Ia explode within a relatively narrow range of delay times.
More recent studies have recovered broader DTD functions, with many converging on a declining power-law of $\sim t^{-1}$ \citep[e.g.,][]{Maoz2017-CosmicDTD,Castrillo2021-DTD,Wiseman2021-DESRates}, though there is some evidence for a steeper slope in galaxy clusters \citep{Maoz2017-CosmicDTD,FriedmannMaoz2018-ClusterDTD}.
It is especially difficult to constrain the DTD for short delay times \citep{MaozMannucci2012-SNeIaReview} because of the need for SN Ia rates at long lookback times and uncertainties in the age estimates of stellar populations.

@@

The uncertainties in the SN Ia DTD propagate into GCE models. In principle, the observed chemical abundance patterns should therefore contain information about the SN Ia DTD, and by extension their progenitors. A striking feature of the [O/Fe] distribution in the Milky Way is the distinct separation into two components, the high- and low-$\alpha$ sequences [references]. [Introduce other MW disk features].

A few studies have compared different DTDs in one-zone chemical evolution models, but their comparisons to abundance data are limited to the local solar neighborhood \citep[e.g.,][]{Andrews2017-ChemicalEvolution,Palicio2023-AnalyticDTD}. 
[Talk about \citet{Matteucci2009-DTDModels} multi-zone models].
\citet{Poulhazan2018-PrecisionPollution} compared six DTDs in a cosmological smoothed-particle hydrodynamics simulation. They found that models with a prompt SN Ia component produced the highest abundance ratios and the narrowest \aFe distributions at late times, although their simulation was not tuned to reproduce the parameters of the Milky Way, and they did not compare to observed chemical abundances. 
The current era of large spectroscopic surveys such as the Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE;][]{Majewski2017-APOGEE} and the forthcoming Milky Way Mapper \citep{Kollmeier2017-SDSS-V} has made abundances across the Milky Way disk available for comparison to more sophisticated GCE models.

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit some parameters that are unchanged from \citetalias{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone model parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:multizone-results} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:multizone-results} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\uptau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration distance \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\uptau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:multizone-results} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:multizone-results} & Maximum simulation time \\
        n               & 8         & \ref{sec:multizone-results} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:multizone-results} & Maximum radius of star formation \\
        M_{g,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        \dot M_r    & continuous    & \ref{sec:multizone-results} & Recycling rate \citep[][Equation 2]{JohnsonWeinberg2020-Starbursts} \\
        \hline
        % DTD
        R_{\rm Ia}(t)   & Equation \ref{eq:dtd-function}    & \ref{sec:dtd-models}  & delay-time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        N_{\rm Ia}/M_\star  & $2.2\times10^{-3}$ M$_\odot^{-1}$ & \ref{sec:yields}  & SNe Ia per unit mass of stars formed \citep{MaozMannucci2012-SNeIaReview} \\
        \hline
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & \ref{sec:yields}  & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & \ref{sec:yields}  & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & \ref{sec:yields}  & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & \ref{sec:yields}  & SN Ia yield of Fe \\
        \hline
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the inside-out SFR \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the late-burst SFR \\
        \uptau_{\rm rise}             & 2 Gyr     & \ref{sec:sfh} & SFR rise timescale for inside-out and early-burst models \\
        \uptau_{\rm EB}(t)          & Equation \ref{eq:earlyburst-taustar}  & \ref{sec:sfh}   & Time-dependence of the early-burst SFE timescale \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the early-burst IFR \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the two-infall IFR \\
        \hline
        % One-zone parameters
        \uptau_\star                    & 2 Gyr & \ref{sec:onezone-results} & SFE timescale in one-zone models \\
        \eta(R_{\rm gal}=8\,{\rm kpc})  & 2.15  & \ref{sec:onezone-results} & Outflow mass-loading factor at the solar annulus \\
        \uptau_{\rm sfh}(R_{\rm gal}=8\,{\rm kpc})    & 15.1 Gyr  & \ref{sec:sfh} & SFH timescale at the solar annulus \\
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

\subsection{Delay-Time Distributions}
\label{sec:dtd-models}

@@

\begin{figure}
    \centering
    \includegraphics{figures/delay_time_distributions.pdf}
    \caption{Selection of models for the SN Ia delay-time distribution (DTD) used in this paper. All functions are normalized such that $f_{\rm Ia}(t=1\,{\rm Gyr})=1$. The black squares represent the DTD recovered for the SDSS-II sample of SNe Ia by \citet{Maoz2012-SloanIIDTD} at the same scale as the DTDs. The horizontal and vertical error bars indicate the time range and 1$\sigma$ uncertainties, respectively, of each DTD measurement. The colored circles along the horizontal axis indicate the median delay time for each model.}
    \label{fig:dtds}
    \script{delay_time_distributions.py}
\end{figure}

% \begin{deluxetable*}{lcll}
% \tablecaption{Summary of SN Ia DTDs explored in this paper.\label{tab:dtds}}
% \tablehead{
% \colhead{Model} & \colhead{Eq.} & \colhead{Parameters} & \colhead{Similar to}
% }
% \startdata
\begin{table*}
    \centering
    \caption{Summary of SN Ia delay-time distributions (DTDs) explored in this paper.}
    \label{tab:dtds}
    \begin{tabular}{lcll}
        \hline\hline
        \multicolumn{1}{c}{Model} & Eq. & \multicolumn{1}{c}{Parameters} & \multicolumn{1}{c}{Similar to} \\
        \hline
        Two-population  & \ref{eq:prompt-dtd}   & $t_{\rm max}=0.05$ Gyr, $\sigma=0.015$ Gyr, & \citet{Mannucci2006-TwoPopulations} \\
                        &                       & $\uptau_{\rm Ia}=3$ Gyr & \\
        Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.4$                 & \citet[][cluster]{Maoz2017-CosmicDTD}; 
                                                      \citet{Heringer2019-FieldGalaxyDTD}       \\
        Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.1$                 & \citet[][field]{Maoz2017-CosmicDTD}; 
                                                      \citet{Wiseman2021-DESRates}              \\
        Exponential & \ref{eq:exponential-dtd}  & $\uptau_{\rm Ia}=1.5$ Gyr    & \citet[][SD]{Greggio2005-AnalyticalRates};
                                                                      \citet{Schonrich2009-RadialMixing};       \\
                    &                           &                   & \citet{Weinberg2017-ChemicalEquilibrium}  \\
        Exponential & \ref{eq:exponential-dtd}  & $\uptau_{\rm Ia}=3$ Gyr   & --- \\
        Plateau     & \ref{eq:plateau-dtd}  & $W=0.3$ Gyr, $\alpha=-1.1$    & \citet[][CLOSE DD]{Greggio2005-AnalyticalRates} \\
        Plateau     & \ref{eq:plateau-dtd}  & $W=1$ Gyr, $\alpha=-1.1$      & \citet[][WIDE DD]{Greggio2005-AnalyticalRates} \\
        Triple-system   & \ref{eq:triple-dtd}   & $f_{\rm init}=0.05f_{\rm peak}$, $t_{\rm rise}=0.5$ Gyr, & \citet{Rajamuthukumar2023-TripleEvolution} \\
                        &                       & $W=0.5$ Gyr, $\alpha=-1.1$ & \\
        \hline
    \end{tabular}
\end{table*}
% \enddata
% \end{deluxetable*}

@@

% [Intro paragraph, motivation for running one-zone models]
Before running the full multi-zone models, it is useful to understand the effects of the DTD in more idealized conditions. A one-zone model assumes the entire gas reservoir is instantaneously mixed, removing all spatial dependence. This limits the ability to compare to observations across the disk, but it eliminates the complicating factor of stellar migration and better isolates the effects of the nucleosynthesis prescription. In this section, we compare the results from one-zone models which examine various parameters of the DTD while keeping other parameters fixed. We use the outputs of our one-zone models to identify the regions in chemical abundance space which are most sensitive to the DTD.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/feh_df_comparison.pdf}
    \caption{Distributions of [Fe/H] from multi-zone models with various models for the SFH and DTD. Each row presents distributions of stars within a range of absolute midplane distance: $1\leq|z|<2$ kpc (\textit{top}), $0.5\leq|z|<1$ kpc (\textit{middle}), and $0\leq|z|<0.5$ kpc (\textit{bottom}). Within each panel, curves of different color represent the distributions of stars binned by Galactocentric radius $R_{\rm gal}$, from $3\leq R_{\rm gal}<5$ kpc (yellow) to $13\leq R_{\rm gal}<15$ kpc (blue). Each distribution is normalized so the area under the curve is 1, and the vertical scale is consistent across each row. All distributions are convolved with observational uncertainties in APOGEE DR17 (see Table \ref{tab:sample}) and smoothed with a box-car width of 0.2 dex. 
    \textit{Left columns:} comparison between the inside-out and two-infall SFHs; both assume the exponential ($\uptau_{\rm Ia}=1.5$ Gyr) DTD. 
    \textit{Center column:} the distributions from APOGEE DR17 for reference, binned and smoothed similarly.
    \textit{Right columns:} comparison between the power-law ($\alpha=-1.4$) and exponential ($\uptau_{\rm Ia}=3$ Gyr) DTDs with the inside-out SFH. The MDFs in the inner Galaxy show the greatest change between the DTDs.}
    \label{fig:feh-df-comparison}
    \script{feh_df_comparison.py}
\end{figure*}

@@

% The choice of DTD and SFH has a larger effect on the inner regions of the Galaxy than the outer. Inner regions have a higher SFR and more stars are formed at early times than late, so differences in the SN Ia rate have a more pronounced effect on the MDF. For all SFHs except two-infall, the outer Galaxy has a longer SFR timescale than the inner Galaxy, so the MDF in the outer Galaxy is relatively unaffected by the fraction of prompt SNe Ia.
The inner Galaxy MDF is more sensitive to the choice of DTD than the outer Galaxy. The SFH peaks earlier and declines more sharply due to the inside-out formation of the disk. Consequently, SNe Ia often explode when the gas supply is significantly lower than when the progenitors formed. This so-called ``gas-starved ISM'' effect drives a faster increase in metallicity \citep[see analytic demonstration in][]{Weinberg2017-ChemicalEquilibrium}, which ultimately lowers the number of low-metallicity stars. The more extended the DTD, the stronger the effect. The outer disk is less affected by the choice of DTD, though, due to the more extended SFH.

@@

The model combination with the best (lowest) score for the MDF is the two-infall SFH with the exponential ($\uptau_{\rm Ia}=3$ Gyr) DTD. The choice of SFH has a larger effect on the overall score than the DTD, and the best-performing SFH is the two-infall model, followed closely by the early-burst. However, the difference between the best-scoring model and the worst (late-burst SFH with triple-evolution DTD) is fairly small. While there are some {\it quantitative} differences in how the shape of the MDF varies with Galactic region, the {\it qualitative} trends are unaffected by the choice of model SFH or DTD. These trends are primarily driven by the assumption of chemical equilibrium, the abundance gradient, and radial migration (see discussion in section 3.2 of \citetalias{Johnson2021-Migration}).

@@

[Intro, connect to previous paragraph]. The relative prominence of the two peaks varies with $R_{\rm gal}$ and $|z|$, as shown by the right-hand panels of Figures \ref{fig:ofe-df-sfh} and \ref{fig:ofe-df-dtd}. A crucial feature of this bimodality, which is not apparent in the previous figures, is the presence of both sequences at the same [Fe/H]. Figure \ref{fig:ofe-bimodality} compares the [O/Fe] distributions in the solar annulus ($7\leq R_{\rm gal}<9$ kpc and $0\leq|z|<2$ kpc) in two bins of [Fe/H] for select model outputs and APOGEE data. The APOGEE distributions in the bottom-right panel (j) show that the high-$\alpha$ mode is more prominent at lower [Fe/H] and vice versa, but the distributions in both bins are clearly bimodal. Moreover, the ``trough'' at ${\rm [O/Fe]}\approx0.2$ is consistent in each bin.

@@

\begin{itemize}
    \item Our diagnostics consistently favor a DTD with a large fraction of delayed SNe Ia, such as the 3 Gyr exponential, 1 Gyr plateau, or triple system evolution models, over a DTD with a high number of prompt SNe Ia, such as the $t^{-1.4}$ power-law or two-population models. [Discuss the disconnect and/or complement with \citet{Maoz2017-CosmicDTD}, etc.; something about different timescales probed by volumetric surveys vs. us]
